---
title: "ExxonMobil Valuation Model and Sensitivity Analysis"
author: "Syndicate Group 5"
date: "21/11/2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---
### References
ExxonMobil Financial Data has been obtained from ExxonMobil's website

All $ values and Balance sheet figures are presented in millions. Date used for market value is company financial year end 31st December 2018
## ExxonMobil constants
### Operating Margin (EBIT%) 
Taking figures from 2018 ExxonMobil Accounts
Revenue = 290,212
Expenditure before Interest and Tax = 260,025
Margin is therefore:
`r (290212-260025)/290212`
### Cost of Capital 
ExxonMobil WACC is between 8 and 12%. will use a midpoint of 10% for initial analysis, based on the FASB standardised FCFF calculation.
### Reinvestment Rate 
ExxonMobil had operating profit of 290,212, and had Capex and R&D costs of 25,923 and 1,116 respectively. This gives a reinvestment rate of 27,039/290,212 = 0.0931, assuming reinvestment grows in line with revenue.
### Tax Rate 
ExxonMobil has a tax reconcilliation in their published accounts, which lists the effective tax rate as 37%

```{r setup, include=FALSE}
source("CorporateFinanceExxonMobil.R")
set.seed(36912)
library(ggplot2)
library(bbplot)
###Modelconstants is a vector with Operating Margin, tax rate, reinvestment rate and cost of capital are inputs into our model
Modelconstants <- c(0.1040, 0.37, 0.0931, 0.1)
###balancesheet is a vector with the balance sheet figures debt, minority interests, cash and non-operating assets
balancesheet <- c(-37796, -6734, 3042, 51122)
```

## Mid point estimate
This valuation will assume a 10 year horizon, a 2.4% Oil price growth rate year on year and a 1.0% oil demand increase. Cost of Capital, tax rate and reinvestment rate is assumed to be constant. 
2.4% approximates the oil price growth projections of Canada Energy Board 
https://apps.cer-rec.gc.ca/ftrppndc/dflt.aspx?GoCTemplateCulture=en-CA 
1.0% demand growth is taken as reasonable based on this McKinsey report on oil.
https://www.mckinsey.com/solutions/energy-insights/global-oil-supply-demand-outlook-to-2035/~/media/231FB01E4937431B8BA070CC55AA572E.ashx
```{r}
#57.88 is the realised revenue per barrel from 2018 financial statements
oil_prices <- linear_growth_function(57.88, 0.024, 11)
#827m is the number of barrels sold in December 2018
barrels_sold <- linear_growth_function(827, 0.01, 11)
#projected revenue growth of chemical
chemical_revenue_growth <- linear_growth_function(32443, 0.034, 11)
#projected revenue growth of downstream
downstream_revenue_growth <- linear_growth_function(221334, 0.034, 11)

share_price_mid <- wrapped_share_price(oil_prices, barrels_sold, chemical_revenue_growth,
                                       downstream_revenue_growth, Modelconstants, balancesheet, 11, 4237)

share_price_mid_table <- PV_forecast_table(oil_prices, barrels_sold, chemical_revenue_growth, downstream_revenue_growth, 11, Modelconstants)
  
print(share_price_mid_table)
print(share_price_mid)

```
#### Current ExxonMobil Share Price
The ExxonMobil share price as at 31st December 2018 close is 
$68.19
https://ir.exxonmobil.com/historical-price-lookup?8c7bdd83-a726-4a84-b969-494be2477e47%5BXOM%5D%5Bdate_month%5D=12&8c7bdd83-a726-4a84-b969-494be2477e47%5BXOM%5D%5Bdate_day%5D=31&8c7bdd83-a726-4a84-b969-494be2477e47%5BXOM%5D%5Bdate_year%5D=2018&url= 
## High growth rate
This estimate uses the same linear growth rate of 2.4% for oil price, but uses a 5% rate for demand growth. 
```{r}
oil_prices <- linear_growth_function(57.88, 0.024, 11)
barrels_sold <- linear_growth_function(826, 0.05, 11)

share_price_high <- wrapped_share_price(oil_prices, barrels_sold, chemical_revenue_growth, downstream_revenue_growth, Modelconstants, balancesheet, 11, 4237)

print(share_price_high)
```
## Low Growth Rate
This uses a growth rate for oil prices of 0.5%, and a demand growth rate of 1%
```{r}
oil_prices <- linear_growth_function(57.88, 0.005, 11)
barrels_sold <- linear_growth_function(826, 0.01, 11)

share_price_low <- wrapped_share_price(oil_prices, barrels_sold, chemical_revenue_growth, downstream_revenue_growth, Modelconstants, balancesheet, 11, 4237)

print(share_price_low)

```

## Negative Growth
This uses a negative growth rate for oil prices of -1%, and a negative growth rate for demand of -1%
```{r}
oil_prices <- linear_growth_function(57.88, -0.01, 11)
barrels_sold <- linear_growth_function(826, -0.01, 11)

share_price_lowest <- wrapped_share_price(oil_prices, barrels_sold, chemical_revenue_growth, downstream_revenue_growth, Modelconstants, balancesheet, 11, 4237)

print(share_price_lowest)

```

## Varying commodity price
```{r}
oil_prices_growth_rates <- seq(-3, 5, 0.1)
oil_prices_growth_rates <- oil_prices_growth_rates/100

price_outcomes_oil <- rep(0, length(oil_prices_growth_rates))
###Modelconstants is a vector with Operating Margin, tax rate, reinvestment rate and cost of capital are inputs into our model
Modelconstants <- c(0.1040, 0.37, 0.0931, 0.1)

for (i in oil_prices_growth_rates){
  oil_prices_o <- linear_growth_function(57.88, i, 11)
  barrels_sold_o <- linear_growth_function(826, 0.01, 11)
  share_price_outcome_year <- wrapped_share_price(oil_prices_o, barrels_sold_o, chemical_revenue_growth, downstream_revenue_growth, Modelconstants, balancesheet, 11, 4237)
  price_outcomes_oil[which(oil_prices_growth_rates == i)] <- share_price_outcome_year
}

oil_price_sensitivity <- data.frame(oil_prices_growth_rates, price_outcomes_oil)


oil_price_sensitivity_plot <- ggplot(oil_price_sensitivity, 
                                     aes( x = oil_prices_growth_rates,
                                          y = price_outcomes_oil)) +
  geom_line(colour = "#1380A1", size = 1)+
                          geom_hline(yintercept = 0, size = 1, colour="#333333") +
  geom_vline(xintercept = 0, size = 1, colour = '#333333') +
  geom_hline(yintercept = 68.13,
  colour = "red", size = 0.5) + 
  theme_light() +
  labs(title="Oil Price growth rate impact on Share Price")

print(oil_price_sensitivity_plot)
```


## Varying the constants - cost of capital
This looks at the impact of the cost of Capital on the valuation of the company, using the mid point growth rates for oil price and demand.
```{r echo=FALSE}
oil_prices <- linear_growth_function(57.88, 0.024, 11)
barrels_sold <- linear_growth_function(826, 0.01, 11)

cost_of_capital_rates <- seq(7, 12, 0.5)
cost_of_capital_rates <- cost_of_capital_rates/100

price_outcomes <- rep(0, length(cost_of_capital_rates))

for (i in seq_len(length(cost_of_capital_rates))){
  Modelconstants[4] <- cost_of_capital_rates[i]
  share_price_outcome <- wrapped_share_price(oil_prices, barrels_sold, chemical_revenue_growth, downstream_revenue_growth, Modelconstants, balancesheet, 11, 4237)
  price_outcomes[i] <- share_price_outcome
}

cost_of_capital_sensitivity <- data.frame(cost_of_capital_rates, price_outcomes)

CoC_plot <- ggplot(cost_of_capital_sensitivity,
        aes(x = cost_of_capital_rates, y = price_outcomes)) +
        geom_line(colour = "#1380A1", size = 1) +
        geom_hline(yintercept = 0, size = 1, colour="#333333") +
        geom_hline(yintercept = 68.13,
        colour = "red", size = 0.5) + 
        theme_light() +
        labs(title="Model Sensitivity to Cost of Capital")
print(CoC_plot)

finalise_plot(plot_name = CoC_plot,
              source_name = "",
              save_filepath = "CostofCapitalSensitivity.png",
              width_pixels = 640,
              height_pixels = 550)

```

## Varying the constants - Operating Margin
This looks at the impact of Operating Margin on the valuation of ExxonMobil, using the mid point growth rates for oil price and demand.
```{r}
oil_prices <- linear_growth_function(57.88, 0.024, 11)
barrels_sold <- linear_growth_function(826, 0.01, 11)

operating_margin_rates <- seq(5, 20, 0.1)
operating_margin_rates <- operating_margin_rates/100

price_outcomes_margin <- rep(0, length(operating_margin_rates))

for (i in seq_len(length(operating_margin_rates))){
  Modelconstants[1] <- operating_margin_rates[i]
  share_price_outcome <- wrapped_share_price(oil_prices, barrels_sold, chemical_revenue_growth, downstream_revenue_growth, Modelconstants, balancesheet, 11, 4237)
  price_outcomes_margin[i] <- share_price_outcome
}

operating_margin_sensitivity <- data.frame(operating_margin_rates, price_outcomes_margin)

op_margin_plot <- ggplot(operating_margin_sensitivity, 
                         aes(x= operating_margin_rates,
                             y= price_outcomes_margin)) +
                          geom_line(colour = "#1380A1", size = 1)+
                          geom_hline(yintercept = 68.13,
                          colour = "red", size = 0.5) + 
                          geom_hline(yintercept = 0, size = 1, colour="#333333") +
  theme_light() +
  labs(title="Operating Margin Impact on Share Price") 

print(op_margin_plot)

```

